---
title: "Exploration"
output: 
  html_document:
    toc: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#IMPORT LIBRARIES
library(jsonlite)
library(tables) # pretty tables
library(pastecs) #stat.desc


#Create data frame for participant file
participant_file <- "data/final_participants.json"
df_subjects <- fromJSON(participant_file)
#names(df_subjects) #print names of each column in the data frame

#Create factor vectors
EXPLICIT <- factor(df_subjects$explicit, levels = c(1,2,3), labels = c("none", "text/image", "interactive"))
IMPASSE  <- factor(df_subjects$impasse, levels = c(1,2), labels = c("none", "impasse"))
AXIS  <- factor(df_subjects$axis)

df_subjects$tt_min <- df_subjects$triangular_time / 1000 / 60
```

###Introduction
The purpose of this study is to compare the relative efficacy of *explicit* (ie. visually-salient text or images that serve instructional purpose) and *implicit* (ie. task and question structure) scaffolding for an unconventional statistical graph.  We hypothesize that implicit techniques---specifically presenting a task that forms a **mental impasse** for the reader---will be more supportive of correct graph interpretation than explicit graph-reading instructions.  We base this hypothesis on prior observation and experimental work (Fox & Hollan, *in preparation*) which revealed that when presented with an unconventional graph with structural elements (ie. axes and grid) that resembled conventional graphs, learners did not always realize they did not understand how the graph was meant to be read. We propose that prior knoweldge of conventional graph systems acts as a form of "graphical fixedness", that learners must overcome to form correct mental models for new representational systems. 


###Methods
####Design
We utilized a 3 (explicit scaffold: none, text/image, interactive) X 2 (implicit scaffold: none, impasse) between-subjects factorial design, with response time and score (max = 15) as dependent variables.   
```{r results = 'asis', echo = FALSE}
#create table of participants by factors (explicit X impasse) 
# table <- tabular( (IMPASSE+1)~(EXPLICIT+1), data = df_subjects )
# table_options(rowlabeljustification="center", justification="right")
# html(table,options=htmloptions(head=TRUE))
```   



####Participants
  
```{r results = 'asis', echo = FALSE}
#Cout number of participants
nSubjects <- nrow(df_subjects)
nFemale <-sum(df_subjects$sex == "Female")
```     
**`r nSubjects` students ** (``r round(nFemale/nSubjects,digits = 2) * 100`` % female) registered as STEM majors at a public university in the United States participated in exchange for course credit (age: ``r min(df_subjects$age)`` - ``r max(df_subjects$age)`` years).

####Measures
```{r echo = FALSE}
# descriptives<-stat.desc(df_subjects)
```
#####Response Accuracy
```{r results = 'asis', echo = FALSE}
#Create a tables of means for task scores 
table<- tabular( (IMPASSE) ~ (Heading("EXPLICIT Scaffold")*EXPLICIT)
                            *(Heading("")*triangular_score) 
                            *(mean+sd)
                            *Format(digits=2),
                            data=df_subjects)
html(table,options=htmloptions(head=TRUE))
#latex(table)
```

#####Response Latency
```{r results = 'asis', echo = FALSE}
#Create a tables of means for task runtime 
table<- tabular( (IMPASSE) ~ (Heading("EXPLICIT Scaffold")*EXPLICIT)
                            *(Heading("")*tt_min) 
                            *(mean + sd)
                            *Format(digits=2), 
                            data=df_subjects)
html(table,options=htmloptions(head=TRUE))
#latex(table)
```





