---
title: "Data Exploration"
output: html_notebook
---
PREREQUISITE: Import dataset into all_participants from all_participants.csv

```{r echo = FALSE}

#stargazer(subjects_time, type = "html",align=TRUE)

#import libraries
library(ggplot2) #graphing
library(psych)   #statistics
library(pastecs)
library(gridExtra)
library(Hmisc)
library(reshape2) #reformatting data
library(dplyr) #data formatting
library(tables) # pretty tables
library(stargazer) #pretty tables


#define labels for experiment order
order_labels <- c(regular = "linear-first", reverse = "triangular-first")
condition_lables <-c( '0' = "no scaffold", '1' = "'what' text", '2' = "'how' text", '3' = "static image", '4' = "interactive image")

#create copy data of base data frame 
df_participantData <- cbind(all_participants)

#create new column with participant order labels
order <- df_participantData$experiment
order[order == "experiment"] <- "regular"
df_participantData$order <- order

#create data subsets
df_subjects <- filter(df_participantData, condition == 0 | condition == 1 |  condition == 2 |condition == 3 | condition == 4 )
df_experts <- filter(df_participantData, condition == 'e0' | condition == 'e4' ) 
df_regular <- subset(df_participantData, order=="regular")
df_reverse <- subset(df_participantData, order=="reverse")
df_regular_subjects <- subset(df_subjects, order=="regular")
df_reverse_subjects <- subset(df_subjects, order=="reverse")

```
###Explore Participant Totals   

We ran 318 subjects through two variants of the study.  In the first variant (regular), participants (n = 155) completed threee tasks: linear, triangular, and drawing tasks. We ran an additional (n = 163) subjects in a second (reverse) variant, with the order of graphing tasks reversed : triangular, linear, and drawing.   

```{r echo = FALSE} 

```   

```{r results = 'asis', echo = FALSE}
#create table of participants by experiment 
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)
table <- tabular(  (CONDITION+1)~(ORDER+1), data = df_subjects )
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```   
.     
     
In addition, we ran `( n = ??)` expert participants through two conditions, composed of research assistants with more than 20 hours of experience analyzing data with the triangular model graph.   

```{r results = 'asis', echo = FALSE}
#create table of participants by experiment 
ORDER <- factor(df_experts$order)
CONDITION <- factor(df_experts$condition)
table <- tabular(  (CONDITION+1)~(ORDER+1), data = df_experts )
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```
*** 
  
###Explore Runtime  
**1. Total Runtime**  
*What is the distribution of runtime (for the whole study)? What are the min, max, mean & SD? Are the data normally distributed?*   

The mean total runtime was approximately 40.3 minutes, with a standard deviation of 8.6 minutes, with values ranging from 20.9 to 66.1 minutes.
<span style="color:red"> Should the high and low values be excluded?</span> A Shapiro-Wilk test for normality yielded a value of W = 0.99, 0.01 < p < 0.05, suggesting that the distibution is significantly non-normal, however, the S-W test is known to be biased toward non-normalcy for large samples. The distribution does appear to be slightly positively skewed. Inspection of the QQ-plot reveals that the distribution can be treated as normal for this sample size.   
.   

```{r results = 'asis', echo = FALSE}
subjects_time <- select(df_subjects, order, condition, TOTAL_T_M)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)*((Heading("TOTAL Runtime")*TOTAL_T_M)*(Heading("mean time")*mean))*(Heading("")*(ORDER+1))), 
                  data = subjects_time)
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```


```{r echo = FALSE}
#DISPLAY HISTOGRAMS
TotalTimeHist <-ggplot(df_subjects, aes(TOTAL_T_M))
TotalTimeHist + geom_histogram(binwidth=1) + facet_wrap(~order,labeller=labeller(order=order_labels)) + 
  coord_cartesian(xlim=c(0,70)) +
  ggtitle("Histogram of Total Runtime") + 
  labs(x="Time (minutes)", y="Number of Participants") + 
  theme_bw()+
  theme(strip.background = element_blank())+
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=14, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```
```{r echo = FALSE}
TotalTimeBoxplot <-ggplot(df_subjects, aes(experiment,TOTAL_T_M))
TotalTimeBoxplot + geom_boxplot() +
  labs(x="Graph Order", y="Time (minutes)")+
  ggtitle("Boxplot of Total Runtime") +
  coord_cartesian(ylim=c(0,70)) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=16, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12)) +
  theme(panel.border = element_blank()) + 
  scale_x_discrete(breaks=c("experiment", "reverse"),labels=c("linear-first", "triangular-first"))
```

```{r echo = FALSE}
#SUBSET DATA BY graphOrder (order)

qplotreg_time <- qplot(sample = df_regular$TOTAL_T_M, main="Q-Q Plot: Total Runtime (LM -first)") + theme_bw() 
qplotrev_time <- qplot(sample = df_reverse$TOTAL_T_M, main="Q-Q Plot: Total Runtime (TM-first)") + theme_bw() 
grid.arrange(qplotreg_time, qplotrev_time, ncol=2)
#HOW TO DO LAYOUT
#lay <- c(1,1,2)
#grid.arrange( plot1, plot2, plot3,plot4, layout_matrix = rbind(c(1,1),c(2,2),c(3,4)))
```

```{r echo = FALSE}
#DISPLAY DESCRIPTIVES
describe(df_subjects$TOTAL_T_M)
stat.desc(df_subjects$TOTAL_T_M, basic=FALSE, norm = TRUE)
#shapiro.test(subjects$TOTAL_T_M)
```

***  
**2. LINEAR TASK Runtime by Graph Order**  
*What is the distribution of runtime for just the linear model task for each level of the graph-order factor? What are the min, max, mean & SD?*

The mean runtime for the linear model task is 9.2 minutes, with a standard deviation of 2.5 minutes, and values ranging from 3.9 to 23.0 minutes. <span style="color:red"> Participant with value of 3.9 minutes is suspicious, review for exclusion. </span>


```{r results = 'asis', echo = FALSE}
subjects_time <- select(df_subjects, order, condition, LM_T_M)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)*((Heading("Linear Task")*LM_T_M)*(Heading("mean time")*mean))*(Heading("")*(ORDER+1))), 
                  data = subjects_time)
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```


```{r echo = FALSE}
LMTimeHist <-ggplot(df_subjects, aes(LM_T_M))
LMTimeHist + geom_histogram(binwidth=1) + facet_wrap(~order,labeller=labeller(order=order_labels)) + 
  coord_cartesian(xlim=c(0,30)) +
  ggtitle("Histogram of Linear Task Runtime") + 
  labs(x="Time (minutes)", y="Number of Participants") + 
  theme_bw() +
  theme(strip.background = element_blank())+
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```
```{r echo = FALSE}
LMTimeBoxplot <-ggplot(df_subjects, aes(order,LM_T_M))
LMTimeBoxplot + geom_boxplot() +
  labs(x="Graph Order", y="Time (minutes)")+
  ggtitle("Boxplot of Linear Task Runtime") +
  coord_cartesian(ylim=c(0,30)) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())+ 
  scale_x_discrete(breaks=c("experiment", "reverse"),labels=c("linear-first", "triangular-first"))
```

```{r echo = FALSE}
describe(df_subjects$LM_T_M)
stat.desc(df_subjects$LM_T_M, basic=FALSE, norm = TRUE)
by(data=df_subjects$LM_T_M, INDICES=df_subjects$order, FUN=describe)
```


***  
**3. TRIANGULAR TASK Runtime by Graph Order**  
*What is the distribution of runtime (for just the triangular model task) for each level of the graph-order factor?  What are the min, max, mean & SD?  Do the data follow a normal distribution?*      

The mean runtime for the linear model task is 10.8 minutes, with a standard deviation of 3.3 minutes, and values ranging from 3.1 to 26.8 minutes. <span style="color:red"> Participant with value of 3.1 minutes is suspicious, review for exclusion. </span>

```{r results = 'asis', echo = FALSE}
subjects_time <- select(df_subjects, order, condition, TM_T_M)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)*((Heading("Triangular Task")*TM_T_M)*(Heading("mean time")*mean))*(Heading("")*(ORDER+1))), 
                  data = subjects_time)
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```


```{r echo = FALSE}
TMTimeHist <-ggplot(df_subjects, aes(TM_T_M))
TMTimeHist + geom_histogram(binwidth=1) + facet_wrap(~order,labeller=labeller(order=order_labels)) + 
  coord_cartesian(xlim=c(0,30)) +
  ggtitle("Histogram of Triangular Task Runtime") + 
  labs(x="Time (minutes)", y="Number of Participants") + 
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```
```{r echo = FALSE}
TMTimeBoxplot <-ggplot(df_subjects, aes(experiment,TM_T_M))
TMTimeBoxplot + geom_boxplot() +
  labs(x="Graph Order", y="Time (minutes)")+
  ggtitle("Boxplot of Triangular Task Runtime") +
  coord_cartesian(ylim=c(0,30)) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())+
  scale_x_discrete(breaks=c("experiment", "reverse"),labels=c("linear-first", "triangular-first"))
```
```{r echo = FALSE}
describe(df_subjects$TM_T_M)
stat.desc(df_subjects$TM_T_M)
by(data=df_subjects$TM_T_M, INDICES=df_subjects$order, FUN=describe)
```
***

**4. DRAWING TASK Runtime by Graph Order**  
*What is the distribution of runtime (for just the linear model task) for each level of the graph-order factor? What are the min, max, mean & SD? Do the data follow a normal distribution?*   

The mean runtime for the drawing task is 13.9 minutes, with a standard deviation of 4.6 minutes, and values ranging from 2.2 to 32.1 minutes. <span style="color:red"> Participant with value of 2.2 minutes is suspicious, review for exclusion. </span>

```{r results = 'asis', echo = FALSE}
subjects_time <- select(df_subjects, order, condition, DT_T_M)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)*((Heading("Drawing Task")*DT_T_M)*(Heading("mean time")*mean))*(Heading("")*(ORDER+1))), 
                  data = subjects_time)
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```

```{r echo = FALSE}
DTimeHist <-ggplot(df_subjects, aes(DT_T_M))
DTimeHist + geom_histogram(binwidth=1) + facet_wrap(~order,labeller=labeller(order=order_labels)) + 
  coord_cartesian(xlim=c(0,30)) +
  ggtitle("Histogram of Drawing Task Runtime") + 
  labs(x="Time (minutes)", y="Number of Participants") + 
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```


```{r echo = FALSE}
DTimeBoxplot <-ggplot(df_subjects, aes(order,DT_T_M))
DTimeBoxplot + geom_boxplot() +
  labs(x="Graph Order", y="Time (minutes)")+
  ggtitle("Boxplot of Drawing Task Runtime") +
  coord_cartesian(ylim=c(0,30)) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank()) +
  scale_x_discrete(breaks=c("regular", "reverse"),labels=c("linear-first", "triangular-first"))
```
```{r echo = FALSE}
describe(df_subjects$DT_T_M)
stat.desc(df_subjects$DT_T_M)
by(data=df_subjects$DT_T_M, INDICES=df_subjects$order, FUN=describe)
```

###Summaries Runtime By Graph Order
Regular Order [Linear - Triangular - Drawing Task]
```{r results = 'asis', echo = FALSE}
#summarize times for regular graph order
subjects_time <- select(df_regular, order, condition, LM_T_M, TM_T_M, DT_T_M, TOTAL_T_M)
ORDER <- factor(df_regular$order)
CONDITION <- factor(df_regular$condition)

table <- tabular(  (CONDITION+1)~
                     (Format(digits=3)*
                        ((Heading("Linear Task")*LM_T_M+Heading("Triangular Task")*TM_T_M+Heading("Drawing Task")*DT_T_M+Heading("Total Task")*TOTAL_T_M) *(Heading("mean time")*mean))), 
                  data = df_regular)
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```
Reverse Order [Linear - Triangular - Drawing Task]
```{r results = 'asis', echo = FALSE}
#df_regular <- subset(df_subjects, order=="regular")
#df_reverse <- subset(df_subjects, order=="reverse")

#summarize times for regular graph order
subjects_time <- select(df_reverse, order, condition, LM_T_M, TM_T_M, DT_T_M, TOTAL_T_M)
ORDER <- factor(df_reverse$order)
CONDITION <- factor(df_reverse$condition)

table <- tabular(  (CONDITION+1)~
                     (Format(digits=3)*
                        ((Heading("Linear Task")*LM_T_M+Heading("Triangular Task")*TM_T_M+Heading("Drawing Task")*DT_T_M+Heading("Total Task")*TOTAL_T_M) *(Heading("mean time")*mean))), 
                  data = df_reverse)
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```


###Summaries Score By Graph Order
Regular Order [Linear - Triangular - Drawing Task]
```{r results = 'asis', echo = FALSE}
#df_regular <- subset(df_subjects, order=="regular")
#df_reverse <- subset(df_subjects, order=="reverse")

#summarize times for regular graph order
subjects_score <- select(df_regular, order, condition, linear_score, triangular_score, d_n, points, scaled_draw_score, totalScore)
ORDER <- factor(df_regular$order)
CONDITION <- factor(df_regular$condition)

table <- tabular(  (CONDITION+1)~
                     (Format(digits=3)*
                        ((Heading("Linear Task")*linear_score+Heading("Triangular Task")*triangular_score+Heading("Drawing Analysis")*d_n+Heading("Drawing Task")*points+Heading("Drawing Total")*scaled_draw_score+Heading("Total")*totalScore) *(Heading("mean score")*mean))), 
                  data = df_regular)
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```

Reverse Order [Linear - Triangular - Drawing Task]
```{r results = 'asis', echo = FALSE}
#df_regular <- subset(df_subjects, order=="regular")
#df_reverse <- subset(df_subjects, order=="reverse")

#summarize times for regular graph order
subjects_score <- select(df_reverse, order, condition, linear_score, triangular_score, d_n, points, scaled_draw_score, totalScore)
ORDER <- factor(df_reverse$order)
CONDITION <- factor(df_reverse$condition)

table <- tabular(  (CONDITION+1)~
                     (Format(digits=3)*
                        ((Heading("Linear Task")*linear_score+Heading("Triangular Task")*triangular_score+Heading("Drawing Analysis")*d_n+Heading("Drawing Task")*points+Heading("Drawing Total")*scaled_draw_score+Heading("Total")*totalScore) *(Heading("mean score")*mean))), 
                  data = df_reverse)
table_options(rowlabeljustification="center", justification="right")
html(table,options=htmloptions(head=TRUE))
```

###Summaries Drawing Interpretation 
Regular Order [Linear - Triangular - Drawing Task]
```{r results = 'asis', echo = FALSE}
#df_regular <- subset(df_subjects, order=="regular")
#df_reverse <- subset(df_subjects, order=="reverse")

#create contingency table for drawing interpretation
tbl = table(df_regular$condition, df_regular$`Graph Type`)
tbl

chisq.test(tbl) 
```

```{r echo = FALSE}
#bar chart of graph interpretations
g <- ggplot(df_regular_subjects, aes(`Graph Type`))
g + geom_bar() + 
  ggtitle("Frequency of Graph Types in Drawing Task") + 
  labs(x="Graph Type in Drawing ", y="Number of Participants") + 
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```



Reverse Order [Triangular - Linear - Drawing Task]
```{r results = 'asis', echo = FALSE}
#df_regular <- subset(df_subjects, order=="regular")
#df_reverse <- subset(df_subjects, order=="reverse")

#create contingency table for drawing interpretation
tbl = table(df_reverse$condition, df_reverse$`Graph Type`)
tbl

chisq.test(tbl) 
```

```{r echo = FALSE}
#bar chart of graph interpretations
g <- ggplot(df_reverse, aes(`Graph Type`))
g + geom_bar() + 
  ggtitle("Frequency of Graph Types in Drawing Task") + 
  labs(x="Graph Type in Drawing ", y="Number of Participants") + 
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```

###Summaries Drawing Time
```{r echo = FALSE}


g <- ggplot(df_regular_subjects, aes(scaled_draw_score, DT_T_M ))
g + geom_point() + 
  ggtitle("Drawing Time and Drawing Score") + 
  labs(x="Drawing Task Score (points)", y="Drawing Time (minutes)") + 
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```