---
title: "Data Exploration"
output: html_notebook
---
```{r echo = FALSE}
order_labels <- c(experiment = "linear-first", reverse = "triangular-first")
library(ggplot2)
library(psych)
library(pastecs)
library(gridExtra)
library(Hmisc)
library(reshape2)

library(readr)
ALL_ITEMS <- read_csv("~/Sites/RESEARCH/2YP-Analysis/xls-files/master-wrangled-data/all_items.csv", 
    col_types = cols(experiment = col_factor(levels = c("experiment", 
        "reverse")), graph = col_factor(levels = c("linear", 
        "triangular")), scenario = col_factor(levels = c("axis", 
        "longmire"))))
```
  
  
  
  
  
  
  
  
  
  
  
  
  
###TITLE
**1. TITLE**  
1. What is the distribution of runtime (for the whole study)?    
2. What are the min, max, mean & SD?   
3. Do they follow a normal distribution?  




```{r echo = FALSE}
#DISPLAY DESCRIPTIVES
describe(ALL_ITEMS$correct)
stat.desc(ALL_ITEMS$correct, basic=FALSE, norm = TRUE)

```

**Analysis**  
<span style="color:#0F5B78">
</span>

<span style = "color:red">
</span>

***  






```{r echo = FALSE}
#DISPLAY DESCRIPTIVES
describe(ALL_ITEMS$correct)
by(data=ALL_ITEMS$correct, INDICES=ALL_ITEMS$condition, FUN=describe)
```

**Analysis**  
<span style="color:#0F5B78">
*The total runtime distributions are similiar for each level of the graph order factor, with the shortest times around 20 minutes and the longest more than 60.*
</span>  

***  
**3. LINEAR TASK Runtime by Graph Order**  
1. What is the distribution of runtime (for just the linear model task) for each level of the graph-order factor?    
2. What are the min, max, mean & SD?   
3. Do they follow a normal distribution?  

```{r echo = FALSE}
LMTimeHist <-ggplot(participantData, aes(LM_T_M))
LMTimeHist + geom_histogram(binwidth=1) + facet_wrap(~experiment,labeller=labeller(order=order_labels)) + 
  coord_cartesian(xlim=c(0,30)) +
  ggtitle("Histogram of Linear Task Runtime") + 
  labs(x="Time (minutes)", y="Number of Participants") + 
  theme_bw() +
  theme(strip.background = element_blank())+
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```
```{r echo = FALSE}
LMTimeBoxplot <-ggplot(participantData, aes(experiment,LM_T_M))
LMTimeBoxplot + geom_boxplot() +
  labs(x="Graph Order", y="Time (minutes)")+
  ggtitle("Boxplot of Linear Task Runtime") +
  coord_cartesian(ylim=c(0,30)) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())+ 
  scale_x_discrete(breaks=c("experiment", "reverse"),labels=c("linear-first", "triangular-first"))
```

```{r echo = FALSE}
describe(participantData$LM_T_M)
by(data=participantData$LM_T_M, INDICES=participantData$experiment, FUN=describe)
```

**Analysis**  
<span style="color:#0F5B78">
*...*
</span>  

***  
**4. TRIANGULAR TASK Runtime by Graph Order**  
1. What is the distribution of runtime (for just the triangular model task) for each level of the graph-order factor?    
2. What are the min, max, mean & SD?   
3. Do they follow a normal distribution?   



```{r echo = FALSE}
TMTimeHist <-ggplot(participantData, aes(TM_T_M))
TMTimeHist + geom_histogram(binwidth=1) + facet_wrap(~experiment,labeller=labeller(order=order_labels)) + 
  coord_cartesian(xlim=c(0,30)) +
  ggtitle("Histogram of Triangular Task Runtime") + 
  labs(x="Time (minutes)", y="Number of Participants") + 
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```
```{r echo = FALSE}
TMTimeBoxplot <-ggplot(participantData, aes(experiment,TM_T_M))
TMTimeBoxplot + geom_boxplot() +
  labs(x="Graph Order", y="Time (minutes)")+
  ggtitle("Boxplot of Triangular Task Runtime") +
  coord_cartesian(ylim=c(0,30)) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())+
  scale_x_discrete(breaks=c("experiment", "reverse"),labels=c("linear-first", "triangular-first"))
```
```{r echo = FALSE}
describe(participantData$TM_T_M)
by(data=participantData$TM_T_M, INDICES=participantData$experiment, FUN=describe)
```


**Analysis**  
<span style="color:#0F5B78">
*...*
</span>  

***  
**5. DRAWING TASK Runtime by Graph Order**  
1. What is the distribution of runtime (for just the linear model task) for each level of the graph-order factor?    
2. What are the min, max, mean & SD?   
3. Do they follow a normal distribution?   

```{r echo = FALSE}
DTimeHist <-ggplot(participantData, aes(D_T_M))
TMTimeHist + geom_histogram(binwidth=1) + facet_wrap(~experiment,labeller=labeller(order=order_labels)) + 
  coord_cartesian(xlim=c(0,30)) +
  ggtitle("Histogram of Drawing Task Runtime") + 
  labs(x="Time (minutes)", y="Number of Participants") + 
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```


```{r echo = FALSE}
TMTimeBoxplot <-ggplot(participantData, aes(experiment,D_T_M))
TMTimeBoxplot + geom_boxplot() +
  labs(x="Graph Order", y="Time (minutes)")+
  ggtitle("Boxplot of Drawing Task Runtime") +
  coord_cartesian(ylim=c(0,30)) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank()) +
  scale_x_discrete(breaks=c("experiment", "reverse"),labels=c("linear-first", "triangular-first"))
```
```{r echo = FALSE}
describe(participantData$D_T_M)
by(data=participantData$D_T_M, INDICES=participantData$experiment, FUN=describe)
```
**Analysis**  
<span style="color:#0F5B78">
*...*
</span>  

***  

**6. Runtime distribution among tasks by Person**  
1. 
2.
3. 



```{r echo = FALSE}
#CREATE STACKED BAR CHART OF RUNTIMES BY TASK
#create subset dataframes [wide] of all time variables

#w.exp.times <- subset(participantData, experiment == "experiment", select = c("subject","LM_T_M","TM_T_M","D_T_M"))
#w.exp.times <-w.exp.times[order(w.exp.times$LM_T_M),]

#m.exp.times <- melt(w.exp.times, id=c("subject"),measured= c("LM_T_M","D_T_M","TM_T_M"))

#ggplot(m.exp.times, aes(x=subject, fill=variable, y=value)) + geom_bar(stat = "identity") + coord_flip() 


```

```{r echo = FALSE}
#LINE GRAPH of MEANS 
#line <- ggplot(participantData, aes(order, TOTAL_T_M))
#line + coord_cartesian(ylim=c(0,70)) +
#theme_bw() + 
#stat_summary(fun.y = mean, geom = "point") + #the point
#stat_summary(fun.y = mean, geom = "line", aes(group=1), color = "blue", linetype="dashed") + #the grouping line
#stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1)
```
When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).
