---
title: "Hypothesis Testing"
output: html_notebook
---
```{r echo = FALSE}
library(ggplot2)
library(psych)
library(pastecs)
library(gridExtra)
library(Hmisc)
library(reshape2)
```

##Study 1: Regular Order (Linear Model - then - Triangular Model)
### H3: (replication) The triangular model times will be LESS than linear model scores, because the triangular model is a more efficient representation for reasoning about temporal intervals (Qiang et. al. 2014). 
<span style = "color: green"> 
*interpretation* 
</span>   

#### a) Are the differences between triangular and linear model response times scores significantly different between scaffold groups? 
> <span style="color:green">   </span>  

```{r echo = FALSE}
w.exp <- subset(participantData, experiment == "experiment")

##construct MEANS PLOT 
line <- ggplot(w.exp, aes(w.exp$condition, w.exp$DIFF_T_M)) 
line + coord_cartesian(ylim=c(-5,5)) +
stat_summary(fun.y = mean, geom = "point") +  #the point
stat_summary(fun.y = mean, geom = "line", aes(group=1), color = "blue", linetype="dashed") + #the grouping line
stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = .1) +
theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
labs(x="scaffold condition", y="time (in minutes)") +
ggtitle("TM Response Time Improvement by Scaffold Condition") +
theme_bw()
```

```{r echo = FALSE}
DiffScoreBoxplot <-ggplot(w.exp, aes(group=condition, x=condition,y=DIFF_T_M))
DiffScoreBoxplot + geom_boxplot() +
  coord_cartesian(ylim=c(-5,5)) +
  labs(x="scaffold condition", y="TM Time Improvement (minutes)")+
  ggtitle("Boxplot of TM Time Improvement") +
  theme_bw() +
  theme(strip.background = element_blank()) +
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```

```{r echo = FALSE}
#compute and evaluate difference in response time
by(w.exp$score_diff, w.exp$condition, stat.desc)

#conduct levene test for homogonaeity of variance (should NOT be significant)
leveneTest(w.exp$DIFF_T_M, w.exp$condition, center=median)

exp_difftimeBYconditionModel <- aov(DIFF_T_M~condition, data = w.exp)
summary(exp_difftimeBYconditionModel)
summary.lm(exp_difftimeBYconditionModel)
```
Levene's IS significant... therefore SHOULD run robust ANOVA...
...1 way independent ANOVA is signficant at p < 0.001. F(1,163) = 21.88, p < 0.001. Multiple R-squared = eta-squared = 0.1183 = r-squared. Effect size r therefore = sqrt(0.1183) = 0.34 a medium effect.    In only the interactive depiction group were participants' accuracy on the triangular model task scores superior to their scores on the linear model task (M = 0.02, SD = 6.66).  In the other groups, the linear model scores were higher.  




