---
output:
  html_notebook: default
  html_document: default
---

_Based on Tutorial by Bodo Winter, available at: http://www.bodowinter.com/tutorial/bw_LME_tutorial2.pdf_

```{r}
library(lme4) #linear mixed effects models
```

```{r}
#load Bodo's sample data
#politeness=read.csv("http://www.bodowinter.com/tutorial/politeness_data.csv")

#attitude = polite or informatl
#scenario = the item task
```

First, let's look at the relationship between politeness and pitch by means of a boxplot. It appears that in both cases, females are higher pitched than males, though there is more overlap in frequency between the two attitutes for men.
```{r} 
boxplot(frequency ~ attitude*gender,
col=c("white","lightgray"),politeness) 
```

Now, let's create the model.  We want to predict frequency based on the fixed effects of attitute and sex, and the random effects of subject, and scenario
politeness.model = lmer(frequency ~ attitude + gender (1|subject) + (1|scenario), data=politeness)

```{r}
politeness.model = lmer(frequency ~ attitude +
gender + (1|subject) +
(1|scenario), data=politeness)
summary(politeness.model)

coef(politeness.model)
```