---
title: "2YP - FOX - Analysis "
output: html_notebook
---

```{r echo = FALSE} 
#IMPORT NECESSARY PACKAGES
library(readr) #load csv
library(pastecs) #stat.desc
library(tables) # pretty tables
library(dplyr) #data formatting
library(ez) #mixed anova
library(reshape2) #reformatting data
library(car)   #statistics
library(ggplot2) #graphs!
library(psych) #describeby
library(gridExtra) #arranging plots
```

```{r echo = FALSE}
#LOAD THE PARTICIPANT DATA
all_participants <- read_csv("~/Sites/RESEARCH/2YP-Analysis/xls-files/master-wrangled-data/all_participants.csv", 
    col_types = cols(ARF = col_skip(), Comments = col_skip(), 
        `Graph Description (if other)` = col_skip(), 
        `Graph Title` = col_skip(), `Graph Type` = col_factor(levels = c("Triangular (right angle)", 
            "Triangular", "Linear Model", 
            "Scatterplot", "Other", "Triangular (asymmetric)")), 
        Points = col_skip(), Rater = col_skip(), 
        `Select each item that is labelled in the graph` = col_skip(), 
        `X Axis` = col_skip(), `Y Axis` = col_skip(),
        GENDER = col_factor(levels = c("Male", 
        "Female","Other")),
        condition = col_factor(levels = c("0", 
            "1", "2", "3", "4")), d_t = col_skip(), 
        d_t_2 = col_skip(), experiment = col_factor(levels = c("experiment", 
            "reverse")), linear_time = col_skip(), 
        lm_scenarios = col_factor(levels = c("axis", 
            "longmire")), ls_t = col_skip(), 
        lt_t = col_skip(), time_diff = col_skip(), 
        tm_scenarios = col_factor(levels = c("longmire", 
            "axis")), triangular_time = col_skip(), 
        ts_t = col_skip(), tt_t = col_skip() ))
#View(all_participants)

#LOAD THE EXPERT DATA
all_experts <- read_csv("~/Sites/RESEARCH/2YP-Analysis/xls-files/master-wrangled-data/all_experts.csv", 
    col_types = cols(
        condition = col_factor(levels = c("0", 
            "1", "2", "3", "4")), d_t = col_skip(), 
        d_t_2 = col_skip(), experiment = col_factor(levels = c("reverse", 
            "experiment")), linear_time = col_skip(), 
        lm_scenarios = col_factor(levels = c("axis", 
            "longmire")), ls_t = col_skip(), 
        lt_t = col_skip(), time_diff = col_skip(), 
        tm_scenarios = col_factor(levels = c("longmire", 
            "axis")), triangular_time = col_skip(), 
        ts_t = col_skip(), tt_t = col_skip()))

```

```{r echo = FALSE}
#SETUP DATA FRAMES
#create copy data of base data frame 
df_subjects <- cbind(all_participants)
df_experts <- cbind(all_experts)

#create a new column for order with better name
df_subjects$order = df_subjects$experiment
levels(df_subjects$order)[levels(df_subjects$order)=="experiment"] <- "LMFirst"
levels(df_subjects$order)[levels(df_subjects$order)=="reverse"] <- "TMFirst"

#create a column for the sum of the two graphing tasks (graphScore)
df_subjects$graphScore = df_subjects$linear_score + df_subjects$triangular_score

#create a column for the sum of the two drawing tasks (graphScore)
df_subjects$drawScore = df_subjects$points + df_subjects$d_n


#create data subsets
df_regular <- subset(df_subjects, experiment=="experiment")
df_reverse <- subset(df_subjects, experiment=="reverse")
```

```{r echo = FALSE}
#SETUP LABELS AND COLORS FOR GRAPHS
order_labels <- c(LMFirst = "linear-first", TMFirst = "triangular-first")
condition_lables <-c( '0' = "no scaffold", '1' = "what-text", '2' = "how-text", '3' = "static image", '4' = "interactive image")
task_labels <- c("LM_T_M" = "linear graph", "TM_T_M" = "triangular graph")

  red = "#F8766D"
  green = "#1cc07e"
  blue = "#59A1FC"
  purple = "#C680FD"
  none = "#7E7D7E"
  text1 = "#071FD5"
  text2 = "#0C91DF"
  img1 = "#EF007A"
  img2 = "#c481d5"

```

```{r echo = FALSE}
#create subsets of runtimes and scores
w_times = subset(df_subjects, select = c("subject","condition", "order", "lm_scenarios", "TM_T_M","LM_T_M","DT_T_M")) #create a wide form subset 
l_times <-melt(w_times, id=c("subject","condition", "order", "lm_scenarios"), measured=c("LM_T_M","TM_T_M","DT_T_M")) #create a longform version

w_scores = subset(df_subjects, select = c("subject","condition", "order", "lm_scenarios", "linear_score","triangular_score","scaled_draw_score")) #create a wide form subset 
l_scores <-melt(w_scores, id=c("subject","condition", "order", "lm_scenarios"), measured=c("linear_score","triangular_score","scaled_draw_score")) #create a longform version
```

##GRAPH RUNTIMES
####Total Runtime by Scaffold
```{r results = 'asis', echo = FALSE}
#MEANS TABLE FOR TOTAL RUNTIME
subjects_total_time <- select(df_subjects, order, condition, TOTAL_T_M)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)
                                    *((Heading("")*TOTAL_T_M)*
                                       (mean + sd))), 
                                      data = subjects_total_time)
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=TRUE))
#latex(table)
#
#print("<H1>Total Time by Scaffold: REGULAR ORDER</H1>")
#describeBy(df_regular$TOTAL_T_M, df_regular$condition)
#print("Total Time by Scaffold: REVERSE ORDER")
#describeBy(df_reverse$TOTAL_T_M, df_reverse$condition)
```
***

####Task Runtime by Scaffold and Graph
```{r results = 'asis', echo = FALSE}
#MEANS TABLE FOR TOTAL RUNTIME
subjects_total_time <- select(df_subjects, order, condition, LM_T_M, TM_T_M)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)
                                    *(Heading("LM")*LM_T_M+Heading("TM")*TM_T_M)*(mean+sd)),  
                                      data = subjects_total_time)
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=TRUE))
#latex(table)
#
#print("<H1>Total Time by Scaffold: REGULAR ORDER</H1>")
#describeBy(df_regular$TOTAL_T_M, df_regular$condition)
#print("Total Time by Scaffold: REVERSE ORDER")
#describeBy(df_reverse$TOTAL_T_M, df_reverse$condition)
```
***

####Task Runtime by Scaffold, Graph and Order
```{r results = 'asis', echo = FALSE}
#MEANS TABLE FOR Time by Task and Order :)
subjects_total_time <- select(df_subjects, order, condition, LM_T_M, TM_T_M)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION)~(Format(digits=3)
                                    *((Heading("Graph Order")*ORDER))
                                    *((Heading("LM")*LM_T_M) + (Heading("TM")*TM_T_M)) 
                                    *(mean + sd)),  
                                      data = subjects_total_time)
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=FALSE))
#latex(table)
#describeBy(df_subjects$LM_T_M, df_subjects$condition)
#describeBy(df_subjects$TM_T_M, df_subjects$condition)
```
***


##GRAPH SCORES
####Total Graph Score by Scaffold
```{r results = 'asis', echo = FALSE}
#MEANS TABLE FOR TOTAL RUNTIME
subjects_total_score <- select(df_subjects, order, condition, graphScore)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)
                                    *((Heading("")*graphScore)*
                                       (mean + sd))), 
                                      data = subjects_total_score)
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=TRUE))
#latex(table)

```
***

####Task Scores by Scaffold and Graph
```{r results = 'asis', echo = FALSE}
#MEANS TABLE FOR TOTAL RUNTIME
subjects_total_score <- select(df_subjects, order, condition, linear_score, triangular_score)
subjects_total_score$linPercent <- round(subjects_total_score$linear_score/15, digits = 1)
subjects_total_score$triPercent <- round(subjects_total_score$triangular_score/15, digits = 1)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)
                                    *((Heading("LM")*linPercent) +Heading("TM")*triPercent)*
                                       (mean + sd)), 
                                      data = subjects_total_score)
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=TRUE))
#latex(table)

```
***

####Task Scores by Scaffold and Graph and Order
```{r results = 'asis', echo = FALSE}
#MEANS TABLE FOR Time by Task and Order :)
subjects_total_score <- select(df_subjects, order, condition, linear_score, triangular_score)
subjects_total_score$linPercent <- round(subjects_total_score$linear_score/15, digits = 1)
subjects_total_score$triPercent <- round(subjects_total_score$triangular_score/15, digits = 1)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION)~(Format(digits=3)
                                    *((Heading("Graph Order")*ORDER))
                                    *((Heading("LM")*linPercent) + (Heading("TM")*triPercent)) 
                                    *(mean + sd)),  
                                      data = subjects_total_score)
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=FALSE))
#latex(table)
#describeBy(df_subjects$LM_T_M, df_subjects$condition)
#describeBy(df_subjects$TM_T_M, df_subjects$condition)
```
*** 

##DRAWING TIME

####Drawing Time by Scaffold

```{r results = 'asis', echo = FALSE}
#MEANS TABLE FOR TOTAL RUNTIME
subjects_draw_time <- select(df_subjects, order, condition, DT_T_M)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)
                                    *((Heading("")*DT_T_M)*
                                       (mean + sd))), 
                                      data = subjects_draw_time)
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=TRUE))
#latex(table)
#
#print("<H1>Total Time by Scaffold: REGULAR ORDER</H1>")
#describeBy(df_regular$TOTAL_T_M, df_regular$condition)
#print("Total Time by Scaffold: REVERSE ORDER")
#describeBy(df_reverse$TOTAL_T_M, df_reverse$condition)
```
***

####Drawing Time by Scaffold and Order
```{r results = 'asis', echo = FALSE}
#MEANS TABLE for drawing time by Scaffold and order
subjects_draw_time <- select(df_subjects, order, condition, DT_T_M)
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)
                                    *(Heading("Task Order")*ORDER)
                                    *((Heading("")*DT_T_M)*
                                       (mean + sd))), 
                                      data = subjects_draw_time)
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=TRUE))
#latex(table)
#
#print("<H1>Total Time by Scaffold: REGULAR ORDER</H1>")
#describeBy(df_regular$TOTAL_T_M, df_regular$condition)
#print("Total Time by Scaffold: REVERSE ORDER")
#describeBy(df_reverse$TOTAL_T_M, df_reverse$condition)
```
***

##DRAWING SCORE

####Drawing Score by Scaffold
```{r results = 'asis', echo = FALSE}
#MEANS TABLE FOR DRAWING SCORE BY SCAFFOLD
subjects_draw_score <- select(df_subjects, order, condition, drawScore)
subjects_draw_score$drawPercent <- subjects_draw_score$drawScore/12
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)
                                    *(Heading("")*drawPercent)*(mean+sd)),  
                                      data = subjects_draw_score )
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=TRUE))
#latex(table)
#
#print("<H1>Total Time by Scaffold: REGULAR ORDER</H1>")
#describeBy(df_regular$TOTAL_T_M, df_regular$condition)
#print("Total Time by Scaffold: REVERSE ORDER")
#describeBy(df_reverse$TOTAL_T_M, df_reverse$condition)
```
***
####Drawing Score by Scaffold and Order
```{r results = 'asis', echo = FALSE}
#MEANS TABLE FOR DRAWING SCORE BY SCAFFOLD
subjects_draw_score <- select(df_subjects, order, condition, drawScore)
subjects_draw_score$drawPercent <- subjects_draw_score$drawScore/12
ORDER <- factor(df_subjects$order)
CONDITION <- factor(df_subjects$condition)

table <- tabular(  (CONDITION+1)~(Format(digits=3)
                                    *(Heading("Order")*ORDER)
                                    *(Heading("")*drawPercent)*(mean+sd)),  
                                      data = subjects_draw_score )
table_options(rowlabeljustification="center", justification="right")
#table
html(table,options=htmloptions(head=TRUE))
#latex(table)
#
#print("<H1>Total Time by Scaffold: REGULAR ORDER</H1>")
#describeBy(df_regular$TOTAL_T_M, df_regular$condition)
#print("Total Time by Scaffold: REVERSE ORDER")
#describeBy(df_reverse$TOTAL_T_M, df_reverse$condition)
```