---
title: "2YP - FOX - Analysis "
author: '-----'
output:
  html_notebook: default
  pdf_document: default
---

```{r echo = FALSE} 
#IMPORT NECESSARY PACKAGES
library(readr) #load csv
library(pastecs) #stat.desc
library(tables) # pretty tables
library(dplyr) #data formatting
library(ez) #mixed anova
library(reshape2) #reformatting data
library(car)   #statistics
library(ggplot2) #graphs!
library(psych) #describeby
library(gridExtra) #arranging plots
```

```{r echo = FALSE}
#LOAD THE PARTICIPANT DATA
all_participants <- read_csv("~/Sites/RESEARCH/2YP-Analysis/xls-files/master-wrangled-data/all_participants.csv", 
    col_types = cols(ARF = col_skip(), Comments = col_skip(), 
        `Graph Description (if other)` = col_skip(), 
        `Graph Title` = col_skip(), `Graph Type` = col_factor(levels = c("Triangular (right angle)", 
            "Triangular", "Linear Model", 
            "Scatterplot", "Other", "Triangular (asymmetric)")), 
        Points = col_skip(), Rater = col_skip(), 
        `Select each item that is labelled in the graph` = col_skip(), 
        `X Axis` = col_skip(), `Y Axis` = col_skip(),
        GENDER = col_factor(levels = c("Male", 
        "Female","Other")),
        condition = col_factor(levels = c("0", 
            "1", "2", "3", "4")), d_t = col_skip(), 
        d_t_2 = col_skip(), experiment = col_factor(levels = c("experiment", 
            "reverse")), linear_time = col_skip(), 
        lm_scenarios = col_factor(levels = c("axis", 
            "longmire")), ls_t = col_skip(), 
        lt_t = col_skip(), time_diff = col_skip(), 
        tm_scenarios = col_factor(levels = c("longmire", 
            "axis")), triangular_time = col_skip(), 
        ts_t = col_skip(), tt_t = col_skip() ))
#View(all_participants)

#LOAD THE EXPERT DATA
all_experts <- read_csv("~/Sites/RESEARCH/2YP-Analysis/xls-files/master-wrangled-data/all_experts.csv", 
    col_types = cols(
        condition = col_factor(levels = c("0", 
            "1", "2", "3", "4")), d_t = col_skip(), 
        d_t_2 = col_skip(), experiment = col_factor(levels = c("reverse", 
            "experiment")), linear_time = col_skip(), 
        lm_scenarios = col_factor(levels = c("axis", 
            "longmire")), ls_t = col_skip(), 
        lt_t = col_skip(), time_diff = col_skip(), 
        tm_scenarios = col_factor(levels = c("longmire", 
            "axis")), triangular_time = col_skip(), 
        ts_t = col_skip(), tt_t = col_skip()))

```

```{r echo = FALSE}
#SETUP DATA FRAMES
#create copy data of base data frame 
df_subjects <- cbind(all_participants)
df_experts <- cbind(all_experts)

#create a new column for order with better name
df_subjects$order = df_subjects$experiment
levels(df_subjects$order)[levels(df_subjects$order)=="experiment"] <- "LMFirst"
levels(df_subjects$order)[levels(df_subjects$order)=="reverse"] <- "TMFirst"

#create a column for the sum of the two graphing tasks (graphScore)
df_subjects$graphScore = df_subjects$linear_score + df_subjects$triangular_score


#create data subsets
df_regular <- subset(df_subjects, experiment=="experiment")
df_reverse <- subset(df_subjects, experiment=="reverse")
```

```{r echo = FALSE}
#SETUP LABELS AND COLORS FOR GRAPHS
order_labels <- c(LMFirst = "linear-first", TMFirst = "triangular-first")
condition_lables <-c( '0' = "no scaffold", '1' = "what-text", '2' = "how-text", '3' = "static image", '4' = "interactive image")
task_labels <- c("LM_T_M" = "linear graph", "TM_T_M" = "triangular graph")

  red = "#F8766D"
  green = "#1cc07e"
  blue = "#59A1FC"
  purple = "#C680FD"
  none = "#7E7D7E"
  text1 = "#071FD5"
  text2 = "#0C91DF"
  img1 = "#EF007A"
  img2 = "#c481d5"

```

```{r echo = FALSE}
#EXPLORE SAMPLE STATISTICS

#record sample size
nSubjects = nrow(df_subjects)
nExperts = nrow(df_experts)
n = nSubjects + nExperts
nReg = nrow(df_regular)
nRev = nrow(df_reverse)

#how many women?
nFemale <-nrow(filter(df_subjects, df_subjects$GENDER == 'Female'))

#describe the sample
descriptives<-stat.desc(df_subjects)
#descriptives

#determine age range
age_min = descriptives$AGE[4]
age_max = descriptives$AGE[5]
age_m = round(descriptives$AGE[9], 2)
age_sd = round(descriptives$AGE[10], 2)


time_min = round(descriptives$TOTAL_T_M[4], 0)
time_max = round(descriptives$TOTAL_T_M[5], 0)
time_m = round(descriptives$TOTAL_T_M[9], 0)
time_sd = round(descriptives$TOTAL_T_M[10], 2)

score_min = round(descriptives$graphScore[4], 0)
score_max = round(descriptives$graphScore[5], 0)
score_m = round(descriptives$graphScore[9], 0)
score_sd = round(descriptives$graphScore[10], 2)

```

> Response Time: Normality? 

_For each of the three tasks, the dependent measures are response time and accuracy (score). Are the response times and scores normally distributed?_

The mean total runtime was approximately `r time_m` minutes, with a standard deviation of `r time_sd` minutes, with values ranging from `r time_min` to `r time_max`  minutes.  
<span style="color:red"> Should the high and low values be excluded?</span>    
A Shapiro-Wilk test for normality yielded a value of W = 0.99, 0.01 < p < 0.05, suggesting that the distibution is significantly non-normal, however, the S-W test is known to be biased toward non-normalcy for large samples. The distribution does appear to be slightly positively skewed. `Inspection of the QQ-plot reveals that the distribution can be treated as normal for this sample size.` <span style="color:red"> Is this a reasonable decision?</span>      

```{r echo = FALSE}
#DISPLAY DESCRIPTIVES
#describe(df_subjects$TOTAL_T_M)

#desc_time <- stat.desc(df_subjects$TOTAL_T_M, basic=FALSE, norm = TRUE)
#desc_time

shapiro.test(df_subjects$TOTAL_T_M)
```

```{r echo = FALSE}
#DISPLAY HISTOGRAMS
TotalTimeHist <-ggplot(df_subjects, aes(TOTAL_T_M)) 
TotalTimeHist + geom_histogram(binwidth=1) + #+ facet_wrap(~order,labeller=labeller(order=order_labels)) + 
  coord_cartesian(xlim=c(0,70)) +
  ggtitle("Histogram of Total Runtime") + 
  labs(x="Time (minutes)", y="Number of Participants") + 
  theme_bw()+
  theme(strip.background = element_blank())+
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=14, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```

```{r echo = FALSE}
#BOXPLOT of TOTAL TIME BY CONDITION & ORDER
BP_totalTime <- ggplot(df_subjects, aes(condition,TOTAL_T_M, fill=order))+ geom_boxplot() +
  labs(x="Scaffold", y="Total Runtime (minutes)", color="Graph Order")+ 
  ggtitle("Total Runtime by Scaffold & Graph Order ") +
  theme_bw() +
  theme(strip.background = element_blank()) +
  scale_fill_manual(   values = c("grey","pink"),
                       name="Graph Order",
                       breaks=c("LMFirst", "TMFirst"),
                       labels=c("linear-first", "triangular-first"))
BP_totalTime

```

```{r echo = FALSE}
#ASSESS NORMALITY OF RUNTIME
##SUBSET DATA BY graphOrder (order)
qplotreg_time <- qplot(sample = df_regular$TOTAL_T_M, main="Q-Q Plot: Total Runtime (LM -first)") + theme_bw() 
qplotrev_time <- qplot(sample = df_reverse$TOTAL_T_M, main="Q-Q Plot: Total Runtime (TM-first)") + theme_bw() 
grid.arrange(qplotreg_time, qplotrev_time, ncol=2)
#HOW TO DO LAYOUT
#lay <- c(1,1,2)
#grid.arrange( plot1, plot2, plot3,plot4, layout_matrix = rbind(c(1,1),c(2,2),c(3,4)))
```
<span style="color:red">Should I exclude the outliers 25 < T > 60 ?</span>

***

> Response Score: Normality? 

The mean total score was approximately `r score_m` points, with a standard deviation of `r score_sd`, with values ranging from `r score_min` to `r score_max`  points.  
<span style="color:red"> 
A Shapiro-Wilk test for normality yielded a value of W = 0.94,  p < 0.001, suggesting that the distibution is significantly non-normal, however, the S-W test is known to be biased toward non-normalcy for large samples. Inspection of the QQ-plot and histogram reveal the distribution of scores is not normal. <span style="color:red"> Is this a reasonable decision?</span>      

```{r echo = FALSE}
#DISPLAY DESCRIPTIVES
#describe(df_subjects$graphScore)

#desc_time <- stat.desc(df_subjects$TOTAL_T_M, basic=FALSE, norm = TRUE)
#desc_time

shapiro.test(df_subjects$graphScore)
```

```{r echo = FALSE}
#DISPLAY HISTOGRAMS
TotalTimeHist <-ggplot(df_subjects, aes(graphScore)) 
TotalTimeHist + geom_histogram(binwidth=1) + #+ facet_wrap(~order,labeller=labeller(order=order_labels)) + 
  coord_cartesian(xlim=c(0,30)) +
  ggtitle("Total Score") + 
  labs(x="Time (minutes)", y="Number of Participants") + 
  theme_bw()+
  theme(strip.background = element_blank())+
  theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=14, hjust=0.5)) + 
  theme(axis.title = element_text(family = "Helvetica", color="#666666", face="bold", size=10)) +
  theme(panel.border = element_blank())
```

```{r echo = FALSE}
#BOXPLOT of TOTAL TIME BY CONDITION & ORDER
BP_totalScore <- ggplot(df_subjects, aes(condition,graphScore, fill=order))+ geom_boxplot() +
  labs(x="Scaffold", y="Total Score (points)", color="Graph Order")+ 
  ggtitle("Total Score by Scaffold & Graph Order ") +
  theme_bw() +
  theme(strip.background = element_blank()) +
  scale_fill_manual(   values = c("grey","pink"),
                       name="Graph Order",
                       breaks=c("LMFirst", "TMFirst"),
                       labels=c("linear-first", "triangular-first"))
BP_totalScore

```

```{r echo = FALSE}
#ASSESS NORMALITY OF RUNTIME

qplotscore <- qplot(sample = df_regular$graphScore, main="Q-Q Plot: Total Score") + theme_bw() 
qplotscore
##SUBSET DATA BY graphOrder (order)
qplotreg_score <- qplot(sample = df_regular$graphScore, main="Q-Q Plot: Total Score (LM -first)") + theme_bw() 
qplotrev_score <- qplot(sample = df_reverse$graphScore, main="Q-Q Plot: Total Score (TM-first)") + theme_bw() 
grid.arrange(qplotreg_score, qplotrev_score, ncol=2)
#HOW TO DO LAYOUT
#lay <- c(1,1,2)
#grid.arrange( plot1, plot2, plot3,plot4, layout_matrix = rbind(c(1,1),c(2,2),c(3,4)))
```










