---
title: "Mixed Effects Analysis"
output: html_notebook
---
*NOTE: Data is imported from all_participants.csv*
```{r echo = FALSE}
#SETUP DATA FRAMES

#import libraries
library(readr) #load csv
library(dplyr) #data formatting
library(ggplot2) #graphing
library(psych)   #statistics
library(pastecs)
#library(gridExtra)
library(Hmisc)
library(reshape2) #reformatting data
#library(tables) # pretty tables
library(ez) #mixed anova
library(nlme) #multilevel model
library(lme4) #better mixed model
library(WRS2) #robust tests
```

```{r echo = FALSE}
#LOAD THE PARTICIPANT DATA
all_participants <- read_csv("~/Sites/RESEARCH/2YP-Analysis/xls-files/master-wrangled-data/all_participants.csv", 
    col_types = cols(ARF = col_skip(), Comments = col_skip(), 
        `Graph Description (if other)` = col_skip(), 
        `Graph Title` = col_skip(), `Graph Type` = col_factor(levels = c("Triangular (right angle)", 
            "Triangular", "Linear Model", 
            "Scatterplot", "Other", "Triangular (asymmetric)")), 
        Points = col_skip(), Rater = col_skip(), 
        `Select each item that is labelled in the graph` = col_skip(), 
        `X Axis` = col_skip(), `Y Axis` = col_skip(),
        GENDER = col_factor(levels = c("Male", 
        "Female","Other")),
        condition = col_factor(levels = c("0", 
            "1", "2", "3", "4")), d_t = col_skip(), 
        d_t_2 = col_skip(), experiment = col_factor(levels = c("experiment", 
            "reverse")), linear_time = col_skip(), 
        lm_scenarios = col_factor(levels = c("axis", 
            "longmire")), ls_t = col_skip(), 
        lt_t = col_skip(), time_diff = col_skip(), 
        tm_scenarios = col_factor(levels = c("longmire", 
            "axis")), triangular_time = col_skip(), 
        ts_t = col_skip(), tt_t = col_skip() ))
#View(all_participants)

#LOAD THE EXPERT DATA
all_experts <- read_csv("~/Sites/RESEARCH/2YP-Analysis/xls-files/master-wrangled-data/all_experts.csv", 
    col_types = cols(
        condition = col_factor(levels = c("0", 
            "1", "2", "3", "4")), d_t = col_skip(), 
        d_t_2 = col_skip(), experiment = col_factor(levels = c("reverse", 
            "experiment")), linear_time = col_skip(), 
        lm_scenarios = col_factor(levels = c("axis", 
            "longmire")), ls_t = col_skip(), 
        lt_t = col_skip(), time_diff = col_skip(), 
        tm_scenarios = col_factor(levels = c("longmire", 
            "axis")), triangular_time = col_skip(), 
        ts_t = col_skip(), tt_t = col_skip()))

```
```{r echo = FALSE}
#SETUP LABELS AND COLORS FOR GRAPHS
order_labels <- c(LMFirst = "linear-first", TMFirst = "triangular-first")
condition_lables <-c( '0' = "no scaffold", '1' = "what-text", '2' = "how-text", '3' = "static image", '4' = "interactive image")
task_labels <- c("LM_T_M" = "linear graph", "TM_T_M" = "triangular graph")

  red = "#F8766D"
  green = "#1cc07e"
  blue = "#59A1FC"
  purple = "#C680FD"
  none = "#7E7D7E"
  text1 = "#071FD5"
  text2 = "#0C91DF"
  img1 = "#EF007A"
  img2 = "#c481d5"

```

```{r echo = FALSE}
#SETUP DATA FRAMES
#create copy data of base data frame 
df_subjects <- cbind(all_participants)
df_experts <- cbind(all_experts)

#create a new column for order with better name
df_subjects$order = df_subjects$experiment
levels(df_subjects$order)[levels(df_subjects$order)=="experiment"] <- "LMFirst"
levels(df_subjects$order)[levels(df_subjects$order)=="reverse"] <- "TMFirst"

#create a column for the sum of the two graphing tasks (graphScore)
df_subjects$graphScore = df_subjects$linear_score + df_subjects$triangular_score

#create a column for the sum of the two drawing tasks (graphScore)
df_subjects$drawScore = df_subjects$points + df_subjects$d_n

#create data subsets
df_regular <- subset(df_subjects, experiment=="experiment")
df_reverse <- subset(df_subjects, experiment=="reverse")
```



####Mixed Designs Analysis
The experimental design consists of 1 within-subjects factor (graph) with two levels (linear, triangular), and two between-subjects factors: scaffold condition (5 levels: no scaffold, 'what' text, 'how' text, static image, interactive image) and task order (linear-first, triangular-first). 

```{r echo = FALSE}
#1. Enter data
# our data df_subjects is in wide format (repeated measures in different columns), however we need it in long format
w_times = subset(df_subjects, select = c("subject","condition", "order", "lm_scenarios", "LM_T_M","TM_T_M")) #create a wide form subset 
l_times <-melt(w_times, id=c("subject","condition", "order", "lm_scenarios"), measured=c("LM_T_M","TM_T_M")) 
#convert other variables to factors
l_times$condition <- factor(l_times$condition)
l_times$order <- factor(l_times$order)
l_times$lm_scenarios <- factor(l_times$lm_scenarios)
```

```{r echo = FALSE}
#2. Explore data
#create boxplot

#SIMPLE BOXPLOT BY CONDITION
#mixedPlot_condition <- ggplot(l_times, aes(condition,value, color=variable))+ geom_boxplot() + 
#  labs(x="scaffold condition", y="time (minutes)") + 
#  ggtitle("Task Response Time by Scaffold ") +
#  theme_bw() 
#mixedPlot_condition
  
#BOXPLOT BY CONDITION, FACET WITH ORDER
mixedPlot_order <- ggplot(l_times, aes(condition,value, fill=variable))+ geom_boxplot() +
  facet_wrap(~order, labeller=labeller(order = order_labels)) +
  labs(x="scaffold condition", y="time (minutes)", color="task")+ 
  ggtitle("Task Response Time by Scaffold & Task Order ") +
  theme_bw() +
  theme(strip.background = element_blank()) +
  scale_fill_manual( values = c(green,red),
                       name="graphing task",
                       breaks=c("LM_T_M", "TM_T_M"),
                       labels=c("linear", "triangular"))
  #use scale_fill_manual to manually change colors 
                      
mixedPlot_order

#BOXPLOT BY CONDITION, FACET WITH ORDER AND SCENARIO
#mixedPlot_scenario <- ggplot(l_times, aes(condition,value, color=variable))+ geom_boxplot() + 
#  facet_grid(lm_scenarios~order, labeller=labeller(order = order_labels)) +
#  theme_bw() +
#  theme(strip.background = element_blank()) +
#  labs(x="scaffold condition", y="time (minutes)")+ 
#  ggtitle("Task Response Time by Scaffold, Task Order and Scenario ") 
#mixedPlot_scenario
``` 

```{r echo = FALSE}
#explore descriptive statistics based on each IV
#by(l_times$value, list(l_times$condition, l_times$order, l_times$variable), stat.desc, basic=FALSE)
```

3. Construct contrasts
```{r echo = FALSE}
#construct contrasts for ANOVA model
lmVStm <-c(1,-1)
contrasts(l_times$variable) <- cbind(lmVStm)
expVSrev<-c(1,-1)
contrasts(l_times$order) <- cbind(lmVStm)
scaffVSold<-c(-4,1,1,1,1)
contrasts(l_times$condition) <- cbind(scaffVSold)
```
4. Compute main model ANOVA
```{r echo = FALSE}

options(contrasts=c("contr.sum","contr.poly"))

#execute the model
mixedModel = ezANOVA(data = l_times,
                     dv = .(value),
                     wid = .(subject),
                     between = .(condition,order,lm_scenarios),
                     within = .(variable),
                     type = 3,
                     detailed = FALSE)
mixedModel
```
RESULTS suggest a significant main-effect of graph task, and significant interactions between condition & graph, and order and task and scenario and task.  Main effects of condition and order were not significant. 

4. Compute the Model with GLM
```{r echo = FALSE}
#build the baseline model with no predictors other than the intercept
baseline <- lme(value ~ 1, random = ~1|subject/condition/order/variable, data= l_times, method="ML")

#add expected main effects of CONDITION
graphModel <- update(baseline, .~. +variable)
#conditionModel<- update(graphModel, .~. +condition)
orderModel<- update(graphModel, .~. +order)
graph_condition <- update(orderModel, .~. +variable:condition)
condition_order <- update(graph_condition, .~. +condition:order)
graph_order <- update(condition_order, .~. +variable:order)
finalModel <- update(graph_order, .~. +variable:lm_scenarios)

anova(baseline, graphModel, conditionModel, orderModel, graph_condition, condition_order, graph_order, finalModel)
summary(orderModel)
```
Condition, GraphXCondition, GraphXOrder, GraphXScenario


```{r}
plot(finalModel)

```

5. Compute contrasts & post-hoc tests